Lexical Richness as a Lens: Exploring the Influence of Social Class on Authors' Writings

### 2024-03-17 paper/paper.qmd
fixing space

```{r}
#| label: tbl-author-cttr-dist
#| tbl-cap: Authors CTTR Average and Standard Deviation
#| echo: false

@@

@tbl-author-cttr-dist shows the average Corrected Type-Token Ratio (CTTR) and the associated standard deviation for each of the four studied authors.

David Herbert Lawrence and Nancy Mitford both have high average CTTRs at 31.61 and 31.97 respectively, indicating a greater lexical richness in their works. However, Lawrence's works show more variability with a standard deviation of 7.37, as opposed to Mitford's 1.49, suggesting her use of vocabulary is more consistent.

Hugh Garners has a notably lower average CTTR of 16.31, paired with a standard deviation of 2.31, which is indicative of a consistently narrower range of vocabulary. Powell Anthony tops the average CTTR at 34.67 with a standard deviation of 1.64, demonstrating not only the highest lexical richness but also a high consistency in vocabulary usage across his works.

These figures collectively offer a quantitative insight into the diversity and consistency of language used by authors from different social backgrounds.

@@

@tbl-multi-lev-model provides coefficients from a negative binomial regression and a multilevel negative binomial regression, indicating how each author's work differs from the intercept in terms of Corrected Type-Token Ratio (CTTR).

Hugh Garners shows a negative coefficient (-0.658 in Neg Binom and -1.308 in Multilevel Neg Binom), suggesting his CTTR is significantly lower than the baseline set by the intercept, which is consistent with the shortest average word count observed in the previous analysis.

Nancy Mitford's coefficient is slightly above zero (0.021 in Neg Binom and 0.152 in Multilevel Neg Binom), indicating her CTTR is almost at the baseline level. This aligns with her position as neither the lowest nor the highest in terms of lexical richness.

Powell Anthony’s positive coefficient (0.107 in Neg Binom and 0.153 in Multilevel Neg Binom) implies his CTTR is above the baseline, though not as high as David Herbert Lawrence's.

David Herbert Lawrence does not have a coefficient listed here but is included in the distribution figure below.

@@

To gain a more intuitive understanding of the model, we visualized the coefficients from @tbl-multi-lev-model in the form of a distribution figure. @fig-multi-lev-model-by-author, presents the distribution of CTTR estimates for each author, providing a visual representation of the variance around the mean estimate.

Powell Anthony’s distribution is relatively narrow and centered around the mean, suggesting his lexical richness is consistently above the baseline with less variation in CTTR scores. Nancy Mitford's distribution also clusters closely around the mean but shows a slight skew towards higher values in the multilevel negative binomial regression, indicating a modest variability with a tendency for higher CTTR scores.

Hugh Garners' distribution is broader and notably shifted to the left of the baseline in both models, implying a lower CTTR and greater variability in lexical richness compared to the others. David Herbert Lawrence, while not explicitly compared against a coefficient in the table, is represented in the figure with the broadest distribution. This suggests that his CTTR is highly variable, confirming the previous analysis which indicated his works are both lengthy and diverse in terms of lexical richness.

Overall, the @fig-multi-lev-model-by-author corroborates the numerical findings of @tbl-multi-lev-model, offering a visual confirmation of the differences in lexical richness and variability among the authors. It highlights the significant spread in the lexical diversity of Lawrence's works, as well as the relative consistency observed in the writings of Powell Anthony and Nancy Mitford, with Garners' works showing both a lower mean and a higher dispersion.

```{r}
#| label: fig-hugh-cttr-vs-wordcount
#| fig-cap: Comparison Between CTTR and Word Count by Authors
#| echo: false

@@

@fig-hugh-cttr-vs-wordcount presents a scatter plot comparing the CTTR and word count for works by David Herbert Lawrence, Hugh Garners, Nancy Mitford, and Powell Anthony. The plot reveals a pattern that suggests a correlation between the lexical richness of an author's work, as measured by CTTR, and the length of the text, as indicated by word count.

David Herbert Lawrence is a conspicuous outlier in this analysis. His works are scattered across a wide range of both CTTR and word counts, showing that while he has a propensity for longer works, the lexical richness within those works is quite variable. Hugh Garners’ points cluster at the lower end of both word count and CTTR scales, suggesting that his works are shorter and less lexically rich. This could reflect the impact of social and educational limitations on his writing style.

Nancy Mitford's works are positioned in the middle range of the word counts but show a higher CTTR, similar to that of Lawrence, indicating a richer vocabulary that is not necessarily dependent on the length of the text. Powell Anthony stands out with a group of works that not only have high CTTR values but also fall into a moderate range of word counts. His data points suggest a strong consistency in lexical richness that does not fluctuate widely with the length of the text.

Overall, @fig-hugh-cttr-vs-wordcount points to a general trend where longer works tend to have higher CTTRs, but also highlights individual variations among authors. David Herbert Lawrence's variability points to a unique stylistic approach, while the other authors demonstrate a more predictable relationship between the volume of text and lexical diversity, aligning with the anticipated influence of their social class and educational experiences on literary production.

@@

@fig-hugh-cttr-vs-wordcount were created using the imports from `data/analysis_data/cttr_all.csv`. Which has the same creation process as @tbl-authors-cttr.

>>>>

## Average CTTR From Data
```{r}
#| label: tbl-author-cttr-dist
#| tbl-cap: Authors CTTR Average and Standard Deviation
#| echo: false

@@

@tbl-author-cttr-dist shows the average Corrected Type-Token Ratio (CTTR) and the associated standard deviation for each of the four studied authors. 
David Herbert Lawrence and Nancy Mitford both have high average CTTRs at 31.61 and 31.97 respectively, indicating a greater lexical richness in their works. However, Lawrence's works show more variability with a standard deviation of 7.37, as opposed to Mitford's 1.49, suggesting her use of vocabulary is more consistent.
Hugh Garners has a notably lower average CTTR of 16.31, paired with a standard deviation of 2.31, which is indicative of a consistently narrower range of vocabulary. Powell Anthony tops the average CTTR at 34.67 with a standard deviation of 1.64, demonstrating not only the highest lexical richness but also a high consistency in vocabulary usage across his works. 
These figures collectively offer a quantitative insight into the diversity and consistency of language used by authors from different social backgrounds.

## Model Visualization

@@

@tbl-multi-lev-model provides coefficients from a negative binomial regression and a multilevel negative binomial regression, indicating how each author's work differs from the intercept in terms of Corrected Type-Token Ratio (CTTR). 
Hugh Garners shows a negative coefficient (-0.658 in Neg Binom and -1.308 in Multilevel Neg Binom), suggesting his CTTR is significantly lower than the baseline set by the intercept, which is consistent with the shortest average word count observed in the previous analysis. 
Nancy Mitford's coefficient is slightly above zero (0.021 in Neg Binom and 0.152 in Multilevel Neg Binom), indicating her CTTR is almost at the baseline level. This aligns with her position as neither the lowest nor the highest in terms of lexical richness. 
Powell Anthony’s positive coefficient (0.107 in Neg Binom and 0.153 in Multilevel Neg Binom) implies his CTTR is above the baseline, though not as high as David Herbert Lawrence's. 
David Herbert Lawrence does not have a coefficient listed here but is included in the distribution figure below.

@@

To gain a more intuitive understanding of the model, we visualized the coefficients from @tbl-multi-lev-model in the form of a distribution figure. @fig-multi-lev-model-by-author, presents the distribution of CTTR estimates for each author, providing a visual representation of the variance around the mean estimate. 
Powell Anthony’s distribution is relatively narrow and centered around the mean, suggesting his lexical richness is consistently above the baseline with less variation in CTTR scores. Nancy Mitford's distribution also clusters closely around the mean but shows a slight skew towards higher values in the multilevel negative binomial regression, indicating a modest variability with a tendency for higher CTTR scores.

Hugh Garners' distribution is broader and notably shifted to the left of the baseline in both models, implying a lower CTTR and greater variability in lexical richness compared to the others. David Herbert Lawrence, while not explicitly compared against a coefficient in the table, is represented in the figure with the broadest distribution. This suggests that his CTTR is highly variable, confirming the previous analysis which indicated his works are both lengthy and diverse in terms of lexical richness.
Overall, the @fig-multi-lev-model-by-author corroborates the numerical findings of @tbl-multi-lev-model, offering a visual confirmation of the differences in lexical richness and variability among the authors. It highlights the significant spread in the lexical diversity of Lawrence's works, as well as the relative consistency observed in the writings of Powell Anthony and Nancy Mitford, with Garners' works showing both a lower mean and a higher dispersion.

## Word Count vs. CTTR

```{r}
#| label: fig-cttr-vs-wordcount
#| fig-cap: Comparison Between CTTR and Word Count by Authors
#| echo: false

@@

@fig-cttr-vs-wordcount presents a scatter plot comparing the CTTR and word count for works by David Herbert Lawrence, Hugh Garners, Nancy Mitford, and Powell Anthony. The plot reveals a pattern that suggests a correlation between the lexical richness of an author's work, as measured by CTTR, and the length of the text, as indicated by word count.
David Herbert Lawrence is a conspicuous outlier in this analysis. His works are scattered across a wide range of both CTTR and word counts, showing that while he has a propensity for longer works, the lexical richness within those works is quite variable. Hugh Garners’ points cluster at the lower end of both word count and CTTR scales, suggesting that his works are shorter and less lexically rich. This could reflect the impact of social and educational limitations on his writing style.

Nancy Mitford's works are positioned in the middle range of the word counts but show a higher CTTR, similar to that of Lawrence, indicating a richer vocabulary that is not necessarily dependent on the length of the text. Powell Anthony stands out with a group of works that not only have high CTTR values but also fall into a moderate range of word counts. His data points suggest a strong consistency in lexical richness that does not fluctuate widely with the length of the text.
Overall, @fig-cttr-vs-wordcount points to a general trend where longer works tend to have higher CTTRs, but also highlights individual variations among authors. David Herbert Lawrence's variability points to a unique stylistic approach, while the other authors demonstrate a more predictable relationship between the volume of text and lexical diversity, aligning with the anticipated influence of their social class and educational experiences on literary production.

@@

@fig-cttr-vs-wordcount were created using the imports from `data/analysis_data/cttr_all.csv`. Which has the same creation process as @tbl-authors-cttr.
